---
title: "When the bootstrap fails"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE,message=FALSE}
library(BIOL710AdvBiometry)
library(learnr)
library(learnrhash)
library(tidyverse)
library(gradethis)
library(openintro)
library(infer)
tutorial_options(
  # use gradethis for checking
  exercise.checker = gradethis::grade_learnr
  )
knitr::opts_chunk$set(echo = FALSE)

fructose_measurements <- tibble(measured = c(39.5, 39.3, 40.3))
bootstrap_samples <- fructose_measurements %>%
    rep_sample_n(size = 3, reps = 50, replace = TRUE)
bootstrap_samples_organize <- bootstrap_samples %>%
  group_by(replicate) %>%
  summarise(min = min(measured), middle = sort(measured)[2], max = max(measured))
no_replicate_column <- bootstrap_samples_organize %>% 
  select(-replicate)
duplicate_info <- bootstrap_samples_organize %>%
  select(-replicate) %>%
  group_by_all() %>%
  summarise(count = n())
```

## Load packages

In this lab, we will explore and visualize the data using the **tidyverse** suite of packages as well as the **openintro** package.

Let's load the packages.

```{r, packages, exercise = T}
library(tidyverse)
library(infer)
```


## Small Data

The national metrology (measurement science) institute of Argentina made triplicate determinations of the mass fraction of fructose in honey using
high-performance liquid chromatography, and obtained the following values.

```{r, data, exercise = T}
fructose_measurements <- tibble(measured = c(39.5, 39.3, 40.3))
```

Use R code to get 50 resamplings of three measurements each (with replacement). Name this dataframe `bootstrap_samples`.

```{r dxqjmkiirhzxgmzj, exercise = TRUE}
bootstrap_samples <- ___ %>%
    ___(size = ___, reps = ___, replace = ___)

head(bootstrap_samples)
```

```{r dxqjmkiirhzxgmzj-hint-1}
bootstrap_samples <- fructose_measurements %>%
    rep_sample_n(size = ___, reps = ___, replace = ___)

head(bootstrap_samples)
```

```{r dxqjmkiirhzxgmzj-solution}
bootstrap_samples <- fructose_measurements %>%
    rep_sample_n(size = 3, reps = 50, replace = TRUE)

head(bootstrap_samples)
```

```{r dxqjmkiirhzxgmzj-check}
# check code
gradethis::grade_code()
```

## Unique samples

We want to know how many of these are duplicates. For each replicate I note which number is the minimum, the maximum, and in the middle.

```{r dupl, exercise = T}
bootstrap_samples_organize <- bootstrap_samples %>%
  group_by(replicate) %>%
  summarise(min = min(measured), middle = sort(measured)[2], max = max(measured))

head(bootstrap_samples)
```

Now we can look for unique rows. 

Use an R command to take out the replicate number column (since that will make every row different) and save this new dataframe as `no_replicate_column`. 
```{r vqrkfejzwamyywci, exercise = TRUE}
no_replicate_column <- ___ %>% 
  ___(-___)
```

```{r vqrkfejzwamyywci-hint-1}
no_replicate_column <- ___ %>% 
  select(-___)
```

```{r vqrkfejzwamyywci-solution}
no_replicate_column <- bootstrap_samples_organize %>% 
  select(-replicate)

```

```{r vqrkfejzwamyywci-check}
# check code
gradethis::grade_code()
```


How many unique samples are there? Get this directly, not through `dim`.

```{r jqbfofluzyjhibjn, exercise = TRUE}
unique_samples <- no_replicate_column %>% distinct()

___(unique_samples)
```

```{r jqbfofluzyjhibjn-solution}
nrow(unique_samples)
```

```{r jqbfofluzyjhibjn-check}
# check code
gradethis::grade_code()
```

## Duplication

Now I want to know how many times these are duplicated. 

```{r dupinfo, exercise = T}
duplicate_info <- bootstrap_samples_organize %>%
  select(-replicate) %>%
  group_by_all() %>%
  summarise(count = n())

head(duplicate_info)
```

What is the maximum number of duplications? The minimum? 

```{r iqiihexkehxctjvj, exercise = TRUE}
___(duplicate_info$___)
___(duplicate_info$___)
```


```{r iqiihexkehxctjvj-solution}
max(duplicate_info$count)
min(duplicate_info$count)
```

```{r iqiihexkehxctjvj-check}
# check code
gradethis::grade_code()
```




## Submit checkpoint

```{r context="server"}
learnrhash::encoder_logic(strip_output = T)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(
  ui_before = div(strong("Submit your hash in the form below."), br(), br()),
  ui_after  = learnrhash::iframe_ui(
    src = "https://docs.google.com/forms/d/e/1FAIpQLSfKAR9MMxfAR8EWsCi-U2Ow4SjsejLlAUba14FBsch-gox0Gg/viewform?usp=sf_link",
    width="700px", height= "800px"
  )
)
```


* * *

This work was created by Sara Stoudt and modified by Jenna Ekwealor.

