---
title: "Two-Way ANOVA: Interactions Between Factors"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(learnr)
library(learnrhash)
library(gradethis)
tutorial_options(
  exercise.checker = gradethis::grade_learnr
)
knitr::opts_chunk$set(echo = FALSE)

data("ToothGrowth")
tooth_data <- ToothGrowth
```

## Getting Started

In this pre-lab, you'll explore the **two-way ANOVA**, a method used to test whether two categorical explanatory variables — and their interaction — affect a quantitative response variable.

We’ll use the `ToothGrowth` dataset, which examines tooth length (`len`) in guinea pigs based on two factors:
- **Supplement type** (`supp`): orange juice (`OJ`) vs. vitamin C in capsule form (`VC`)
- **Dose** (`dose`): amount of vitamin C administered in milligrams

The dataset comes from a classic study on vitamin C and tooth growth in guinea pigs, where researchers tested whether delivery method and dosage affected tooth development.

### Load the data

```{r glimpse-tooth, exercise = TRUE}
glimpse(tooth_data)
```

### What variables are we working with?

```{r variable-check}
question("Which of the following is the response variable?",
  answer("dose (supplement amount)"),
  answer("supp (supplement type)"),
  answer("ToothGrowth (dataframe)"),
  answer("len (tooth length)", correct = TRUE),
  allow_retry = TRUE
)
```

### How many levels are there for each factor?

```{r factor-levels, exercise = TRUE}
tooth_data %>%
  count(supp)

tooth_data %>%
  count(dose)
```


## Visualizing the data

Let’s explore how dose and supplement type affect tooth length. Fill in the blanks to complete a grouped boxplot:

```{r boxplot-interaction, exercise = TRUE}
ggplot(tooth_data, aes(x = factor(___), y = ___, fill = ___)) +
  ___(position = position_dodge(0.8)) +
  labs(x = "Dose (mg)", y = "Tooth Length")
```

```{r boxplot-interaction-solution}
ggplot(tooth_data, aes(x = factor(dose), y = len, fill = supp)) +
  geom_boxplot(position = position_dodge(0.8)) +
  labs(x = "Dose (mg)", y = "Tooth Length")
```

```{r boxplot-interaction-check}
gradethis::grade_this_code()
```


## Run a two-way ANOVA

Let’s test whether dose, supplement type, or their interaction affect tooth length.

Fill in the blanks below to complete the two-way ANOVA model. Use the `aov()` function and specify both main effects and the interaction.


```{r twoway-anova, exercise = TRUE}
aov_tooth <- aov(___ ~ ___ * ___, data = tooth_data)
summary(___)
```

```{r twoway-anova-solution}
aov_tooth <- aov(len ~ supp * dose, data = tooth_data)
summary(aov_tooth)
```

```{r twoway-anova-check}
gradethis::grade_this_code()
```

### Interpret the output

*Write your interpretation of the p-values for each term (supp, dose, and supp:dose) below.*

```{r interpretation-twoway, exercise = TRUE}
# What does the ANOVA table tell you?
```


## Visualizing the interaction

An **interaction** occurs when the effect of one factor depends on the level of another factor. In the `ToothGrowth` dataset, this means asking: *Does the effect of vitamin C dosage depend on whether the supplement is given as orange juice or a capsule?*

To visualize this, we plot the **mean tooth length** at each dose, separated by supplement type. If the lines are **not parallel**, this suggests an interaction.

We also move the legend onto the plot area so it’s more visible on all screens.

```{r interaction-plot, exercise = TRUE}
ggplot(tooth_data, aes(x = factor(dose), y = len, color = supp, group = supp)) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Interaction Plot", x = "Dose (mg)", y = "Mean Tooth Length") +
  theme(legend.position = "top")
```



```{r interaction-question}
question("If the lines in an interaction plot are not parallel, what does this suggest?",
  answer("There may be an interaction effect", correct = TRUE),
  answer("The main effect of one variable is stronger"),
  answer("The model is nonlinear"),
  answer("There is no effect of either variable"),
  allow_retry = TRUE
)
```


## Summary Visualization with p-value

Let’s put it all together: create a boxplot of the data and add the global ANOVA p-value.

```{r summary-plot, exercise = TRUE}
aov_tooth <- aov(len ~ supp * dose, data = tooth_data)
p_val <- summary(aov_tooth)[[1]]["Pr(>F)"][1]

ggplot(tooth_data, aes(x = factor(dose), y = len, fill = supp)) +
  geom_boxplot(position = position_dodge(0.8)) +
  labs(
    title = "Tooth Length by Dose and Supplement Type",
    x = "Dose (mg)", y = "Tooth Length"
  ) +
  geom_text(
    aes(x = 0.5, y = 30,
        label = paste("Global ANOVA p-value:", round(p_val, 4))),
    inherit.aes = FALSE
  ) +
  theme(legend.position = "top")
```

## Check your understanding

```{r final-interpretation-question}
question("Which of the following best summarizes the two-way ANOVA results?",
  answer("Dose affects tooth length, and the effect of dose may depend on supplement type.", correct = TRUE),
  answer("Supplement type is the only significant factor, dose has no effect."),
  answer("There is no interaction, so neither factor is significant."),
  answer("Only the interaction term is significant; main effects don't matter."),
  allow_retry = TRUE
)
```

## Submit checkpoint

```{r context="server"}
learnrhash::encoder_logic(strip_output = TRUE)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(
  ui_before = div(strong("Submit your hash in the form below."), br(), br()),
  ui_after  = learnrhash::iframe_ui(
    src = "https://docs.google.com/forms/d/e/1FAIpQLSf0FKvlOHfchtSxwIpD9Jo65sDlTjeokodsO1vP64rLcVYZ2A/viewform?usp=header",
    width = "900px", height = "1000px"
  )
)
```

* * *

This work is licensed under a
<a rel="license" href="http://creativecommons.org/licenses/by-sa/4.0/">
  Creative Commons Attribution-ShareAlike 4.0 International License. 
</a> and adapted for use by your instructor.
