---
title: "One-Way ANOVA: Comparing Group Means"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(learnr)
library(learnrhash)
library(gradethis)
tutorial_options(
  exercise.checker = gradethis::grade_learnr
)
knitr::opts_chunk$set(echo = FALSE)

plant_growth <- PlantGrowth
```

## Getting Started

In this pre-lab, you'll explore the **one-way ANOVA**, used to compare means across more than two groups. Our dataset, `PlantGrowth`, contains weights of plants grown under three different conditions:
- `ctrl`: control
- `trt1`: treatment 1
- `trt2`: treatment 2

### Load the data

```{r glimpse-data, exercise = TRUE}
glimpse(plant_growth)
```

### Which variable is the response?

```{r response-var-question}
question("Which of the following is the response variable?",
  answer("weight (plant weight)", correct = TRUE),
  answer("group (treatment group)"),
  answer("PlantGrowth (dataset name)"),
  allow_retry = TRUE
)
```

### How many plants per group?

```{r count-groups, exercise = TRUE}
plant_growth %>% count(group)
```

## Visualizing the data

```{r boxplot-groups, exercise = TRUE}
ggplot(plant_growth, aes(x = group, y = weight)) +
  geom_boxplot() +
  labs(x = "Group", y = "Plant Weight")
```

## Run a one-way ANOVA

Fill in the blanks to complete the model.

```{r oneway-anova, exercise = TRUE}
aov_plant <- aov(___ ~ ___, data = plant_growth)
summary(___)
```

```{r oneway-anova-solution}
aov_plant <- aov(weight ~ group, data = plant_growth)
summary(aov_plant)
```

```{r oneway-anova-check}
gradethis::grade_this_code()
```

### Interpret the output

```{r oneway-anova-interpretation, exercise = TRUE}
# What does the ANOVA table suggest about differences among group means?
```

## Summary Visualization with p-value

```{r summary-plot-setup, include=FALSE}
aov_plant <- aov(weight ~ group, data = plant_growth)
```

```{r summary-plot, exercise = TRUE}
p_val <- summary(aov_plant)[[1]]["Pr(>F)"][1, 1]

ggplot(plant_growth, aes(x = group, y = weight)) +
  geom_boxplot() +
  labs(
    title = "Plant Weight by Treatment Group",
    x = "Group", y = "Plant Weight"
  ) +
  annotate("text", x = 1, y = 7, label = sprintf("Global ANOVA p-value: %.4f", as.numeric(p_val))) +
  theme(legend.position = "none")
```

### Check your understanding

```{r interpretation-question-final}
question("What does the one-way ANOVA indicate about plant weights?",
  answer("At least one treatment affects plant growth.", correct = TRUE),
  answer("All groups have equal means."),
  answer("There is no variation in the data."),
  answer("The ANOVA model failed."),
  allow_retry = TRUE
)
```

## Submit checkpoint

```{r context="server"}
learnrhash::encoder_logic(strip_output = TRUE)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(
  ui_before = div(strong("Submit your hash in the form below."), br(), br()),
  ui_after  = learnrhash::iframe_ui(
    src = "https://docs.google.com/forms/d/e/1FAIpQLSf0FKvlOHfchtSxwIpD9Jo65sDlTjeokodsO1vP64rLcVYZ2A/viewform?usp=header",
    width = "900px", height = "1000px"
  )
)
```

* * *

This work is licensed under a
<a rel="license" href="http://creativecommons.org/licenses/by-sa/4.0/">
  Creative Commons Attribution-ShareAlike 4.0 International License. 
</a> and adapted for use by your instructor.
