---
title: "Comparing Two Means: t-test"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---
```{r setup, include=FALSE, message=FALSE}
library(BIOL710)
library(learnr)
library(learnrhash)
library(tidyverse)
library(gradethis)
library(infer)
tutorial_options(
  exercise.checker = gradethis::grade_learnr
)
knitr::opts_chunk$set(echo = FALSE)

# Load and filter data
control_cells <- heat_shock_response %>%
  dplyr::filter(condition == "control")
heat_stressed_cells <- heat_shock_response %>%
  dplyr::filter(condition == "heat_stressed")
```

### Getting Started

In this lab, you'll explore the t-test, a statistical method used to compare two independent group means and assess if they are significantly different from each other.

### Load Packages

We will analyze data using the `tidyverse` and `infer` packages. Load the necessary packages below:

```{r}
library(tidyverse)
library(infer)
```

### The Data

We will analyze experimental data measuring heat shock protein (HSP) expression levels in control versus heat-stressed mammalian cells.

```{r}
glimpse(heat_shock_response)
head(heat_shock_response)
```

### Comparing Means with a t-test

The goal is to compare the mean expression levels of HSPs between control and heat-stressed cells.

#### Step 1: Compute Summary Statistics

Calculate the mean and standard deviation of HSP expression for both conditions.

```{r}
control_mean <- mean(control_cells$HSP_expression)
control_sd <- sd(control_cells$HSP_expression)
heat_stressed_mean <- mean(heat_stressed_cells$HSP_expression)
heat_stressed_sd <- sd(heat_stressed_cells$HSP_expression)
```

#### Step 2: Visualizing the Distributions

Create histograms to visualize the distributions of HSP expression levels in both conditions.

```{r}
ggplot(control_cells, aes(x = HSP_expression)) + geom_histogram()
ggplot(heat_stressed_cells, aes(x = HSP_expression)) + geom_histogram()
```

#### Step 3: Conducting a t-test

We will now perform an independent t-test to determine if the mean HSP expression differs significantly between the control and heat-stressed cells.

```{r}
t_test_result <- t.test(control_cells$HSP_expression, heat_stressed_cells$HSP_expression)
print(t_test_result)
```

#### Step 4: Interpretation

Review the t-test output and determine if the p-value suggests a significant difference.

What does the p-value indicate about the difference in HSP expression levels between the two conditions?

*Provide your interpretation here.*

#### Step 5: Visualization with the infer Package

Using the infer package, we can further visualize the hypothesis testing process.

```{r}
heat_shock_response %>%
  specify(response = HSP_expression, explanatory = condition) %>%
  hypothesize(null = "independence") %>%
  calculate(stat = "t")
```

### Summary

In this lab, you learned how to:
- Calculate summary statistics for comparing two groups.
- Visualize distributions.
- Perform an independent t-test.
- Interpret the results.
- Use the infer package for hypothesis testing.

Make sure to review the results and think critically about what they tell us about the data!
